### 2017-07-04 opsem.tex
Edits.

Fig.~\ref{fig:txnimp} shows the syntax and small-step semantics of
\txnimp, a core language that we use to formalize our intuitions about
reasoning under weak isolation. Variables ($x$), Integer and boolean
constants ($k$), records ($r$) of named constants, sets ($s$) of such
records, arithmetic and boolean expressions ($e_1 \odot e_2$), and
record expressions ($\{\bar{f}=\bar{e}\}$) constitute the syntactic
class of expressions ($e$). Commands ($c$) include $\cskip$,
conditional statements, \C{LET} constructs to bind names, \C{FOREACH}
loops, SQL statements, their sequential composition ($c_1;c_2$), and
transactions ($\ctxn{i}{\I}{c}$) and their parallel composition
($c_1\,||\,c_2$). The $\I$ in the \C{TXN} block syntax is the
transaction's isolation specification, which will be explained later.
Certain terms that only appear at run-time are also present in $c$.
These include a \R{TXN} block tagged with sets ($\stl$ and $\stg$) of
records whose meaning is explained below, and a runtime \R{FOREACH}
expression that keeps track of the set ($s_1$) of items already
iterated, and the set ($s_2$) of items yet to be iterated. Note that
even the surface-level syntax of the \C{FOREACH} command shown here is
a little different from the one used in previous sections; its
higher-order function has two arguments, $y$ and $z$, which are
invoked (during the reduction) with the set of already-iterated items,
and the current item, respectively. This form of \C{FOREACH} lends
itself to inductive reasoning that will be useful for verification
(Sec.~\ref{sec:reasoning}).

@@

The top-level reduction relation defines the small-step semantics of
transactions, and their parallel composition. A transaction always
comes tagged with an isolation specification $\I_e$ or $\I_c$ that
dictates the timing and nature of interferences that a transaction can
witness, both during its execution ($\I_e$), and when it is about to
commit ($\I_c$).  Formally, $\I_e$ and $\I_c$ are functions that map
the (current) transaction-local database state ($\stl$), the state
($\stg$) of the global database when the transaction last took a step,
and the current state ($\stg'$) of the global database.  Intuitively,
$\stg'\neq\stg$ indicates an interference from another concurrent
transaction, and the predicates $\I_e$ and $\I_c$ decide if this
interference is allowed or not, taking into account the local database
state ($\stl$). For instance, as described in
(Sec.~\ref{sec:motivation}), an RR transaction on PostgreSQL defines
$\I$ as following:
\begin{smathpar}
\begin{array}{lcl}
\I_e\,\,(\stl,\stg,\stg') & = & \stg' = \stg\\
\I_c\,\,(\stl,\stg,\stg') & = & \forall(r\in\stl)(r'\in\stg).~ r'.\idf = r.\idf \Rightarrow r'\in\stg'.
\end{array}
\end{smathpar}

The reduction of a $\ctxn{i}{\I_e}{c}$ begins by first converting it
to its run-time form $\ctxn{i}{\I_e,\stl,\stg}{c}$, where $\stl =
\emptyset$, and $\stg$ the current (global) database.  Rule
\rulelabel{E-Txn} reduces $\ctxn{i}{\I_e,\stl,\stg}{c}$ under a
database state ($\stg'$), only if the isolation specification ($\I$)
allows the interference between $\stg$ and $\stg'$ during that
step. Rule \rulelabel{E-Commit} commits the transaction
$\ctxn{i}{\I_c,\stl,\stg}{c}$ by flushing its uncommitted records to
the database. This is done only if the interference between $\stg$ and
$\stg'$ is allowed at the commit point by the isolation specification.
The distinction between $\I_e$ and $\I_c$ allows us to model realistic
database features such as snapshot isolation that isolates the effects
a transaction can witness during its execution but exposes
interferences from concurrent transactions at the commit point.

>>>>

Fig.~\ref{fig:txnimp} shows the syntax and small-step semantics of
\txnimp, a core language that we use to formalize our intuitions about
reasoning under weak isolation. Variables ($x$), Integer and boolean
constants ($k$), records ($r$) of named constants, sets ($s$) of such
records, arithmetic and boolean expressions ($e_1 \odot e_2$), and
record expressions ($\{\bar{f}=\bar{e}\}$) constitute the syntactic
class of expressions ($e$). Commands ($c$) include $\cskip$,
conditional statements, \C{LET} constructs to bind names, \C{FOREACH}
loops, SQL statements, their sequential composition ($c_1;c_2$), and
transactions ($\ctxn{i}{\I}{c}$) and their parallel composition
($c_1\,||\,c_2$). The $\I$ in the \C{TXN} block syntax is the
transaction's isolation specification, which will be explained later.
Certain terms that only appear at run-time are also present in $c$.
These include a {\sf txn} block tagged with sets ($\stl$ and $\stg$) of
records whose meaning is explained below, and a runtime {\sf foreach}
expression that keeps track of the set ($s_1$) of items already
iterated, and the set ($s_2$) of items yet to be iterated. Note that
even the surface-level syntax of the \C{FOREACH} command shown here is
a little different from the one used in previous sections; its
higher-order function has two arguments, $y$ and $z$, which are
invoked (during the reduction) with the set of already-iterated items,
and the current item, respectively. This form of \C{FOREACH} lends
itself to inductive reasoning that will be useful for verification
(Sec.~\ref{sec:reasoning}).

@@

The top-level reduction relation defines the small-step semantics of
transactions, and their parallel composition. A transaction always
comes tagged with an isolation specification $\I$, which has two
components $\I_e$ and $\I_c$, the dictate the timing and nature of
interferences that the transaction can witness, during its execution
($\I_e$), and when it is about to commit ($\I_c$).  Formally, $\I_e$
and $\I_c$ are predicates over by the (current) transaction-local
database state ($\stl$), the state ($\stg$) of the global database
when the transaction last took a step, and the current state ($\stg'$)
of the global database.  Intuitively, $\stg'\neq\stg$ indicates an
interference from another concurrent transaction, and the predicates
$\I_e$ and $\I_c$ decide if this interference is allowed or not,
taking into account the local database state ($\stl$). For instance,
as described in (Sec.~\ref{sec:motivation}), an RR transaction on
PostgreSQL defines $\I$ as following:
\begin{smathpar}
\begin{array}{lcl}
\I_e\,\,(\stl,\stg,\stg') & = & \stg' = \stg\\
\I_c\,\,(\stl,\stg,\stg') & = & \forall(r\in\stl)(r'\in\stg).~ r'.\idf = r.\idf \Rightarrow r'\in\stg'.
\end{array}
\end{smathpar}
For brevity in presentation, we use $\I$ instead of $\I_e$ and $\I_c$
whenever the discussion applies to both $\I_e$ and $\I_c$.

The reduction of a $\ctxn{i}{\I}{c}$ begins by first converting it to
its run-time form $\ctxnr{i}{\I,\stl,\stg}{c}$, where $\stl =
\emptyset$, and $\stg$ the current (global) database.  Rule
\rulelabel{E-Txn} reduces $\ctxnr{i}{\I,\stl,\stg}{c}$ under a
database state ($\stg'$), only if the execution-time isolation
specification ($\I_e$) allows the interference between $\stg$ and
$\stg'$. Rule \rulelabel{E-Commit} commits the transaction
$\ctxnr{i}{\I,\stl,\stg}{c}$ by flushing its uncommitted records to
the database. This is done only if the interference between $\stg$ and
$\stg'$ is allowed at the commit point by the isolation specification
($\I_c$).  The distinction between $\I_e$ and $\I_c$ allows us to
model snapshot semantics of realistic isolation levels that isolate a
transaction from interference during its execution, but exposes
interference at the commit point.
